Modeling the impacts of park access on health outcomes: a utility-based accessibility approach

### 2020-09-22 manuscript/manuscript.Rmd
Submit first revisions

---
title: >
  Modeling the impacts of park access on health outcomes: a utility-based 
  accessibility approach
runninghead: Macfarlane \emph{et al}.
author:
  - name: Gregory Macfarlane
    num: 1
  - name: Nico Boyd
    num: 2
  - name: John E. Taylor
    num: 3
  - name: Kari Watkins
    num: 3
address:
  - code: 1
    address: Civil and Environmental Engineering Department, Brigham Young University 
  - code: 2
    address: Fehr and Peers, Los Angeles, California
  - code: 3
    address: School of Civil and Environmental Engineering, Georgia Institute of Technology
corrauth: "Gregory Macfarlane, 430 Engineering Building, Provo UT 84663"
email: gregmacfarlane@gmail.com
abstract: >
  Recent research has underscored the potential for public green spaces to influence individual and societal health outcomes, but empirical measurements of such influences have yielded mixed results to date, with particular disagreement surrounding how access to parks ought to be defined while controlling for alternate explanations. In this paper we apply a comprehensive measure of park accessibility drawn from random utility choice theory, which avoids arbitrary assertions of proximity while incorporating potentially numerous amenities and attributes of both the parks and the population. We apply this utility-based accessibility measure to correlate Census tract-level obesity and physical activity rate estimates from the Centers for Disease Control and Prevention's 500 Cities project with tract-level American Community Survey socioeconomic data in New York City, paired with geographic open space data from New York City. Controlling for the socioeconomic variables and spatially correlated error terms, we show a positive and significant relationship between park access and physical activity rates. The data also suggest a negative relationship between park access and obesity rates, beyond what is expected through physical activity and socioeconomics. In doing so, this research contributes a more comprehensive modeling approach for measuring the impact of park access on health, and may improve our understanding of the role  parks and access to them can serve in furthering public health objectives.
keywords: Accessibility; Green Space
classoption:
  - shortAfour
output: 
  rticles::sage_article:
    keep_md: yes
    keep_tex: yes
    includes: 
      in_header: preamble.tex
      after_body: appendix.tex
bibliography: library.bib
bibliographystyle: sageh.bst
---

@@

Applications of this isochrone-based framework are easily derived and relatively
common. 
ParkScore [@parkscore2019] is a sophisticated application of this approach where $D$ is a
carefully calculated 10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. @devries2003natural
showed the amount of greenspace within a 3-kilometer radius was positively associated
with a suite of self-reported health outcomes. Similarly, @devries2013streetscape
showed that streetscape greenery within 500 meters of a residence was associated with
perceived general health. Conversely, @carlson2012interactions found no relationship
between the health of senior citizens and the number of parks within 500 meters.

@@

In a destination choice framework, the tracts represent the "origins" and the
"destinations" are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these distinct sub-facilities, we created 
variables for each containing park indicating the presence of sports courts,
playgrounds, and trails.

@@

Note that in the lag-$X$ specifications (SLX and SDEM) we exclude the accessiblity 
component $\boldsymbol{A}$, that is we do *not* consider that the accessibility
in a neighboring zone will have an effect on physical activity or obesity rates.
In a practical sense, this is because the accessibility term $\boldsymbol{A}$ is itself spatially
determined.

@@

To determine which spatial effects specification was appropriate, we estimated
SLX and SDEM models of physical activity participation rates as a function of
tract covariates and  accessibility. These two models are presented in Table 
\ref{tab:slx-sdem}. With only a few exceptions, the SLX and SDEM coefficient
estimates for corresponding covariates lie within the 95\% confidence intervals
of the other model. A Hausman-style test [@Pace2008] on the coefficients 
assumes a null hypothesis that the coefficients are equivalent; this test 
produces $p$-value of only $`r slx_sdem_htest[2]`$. Though this statistic is
well beneath a common hypothesis test threshold of $\alpha = 0.05$, we
elect retain the null hypothesis that the coefficients are equivalent, as 
the differences are at no point practically different from each other. The consequences
of rejecting the null hypothesis in this case would be to suggest a model with 
spatial dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we adopt the
SDEM specification moving forward as the spatial lag of the errors is 
significant and the overal log likelihood of the SDEM is higher than the SLX 
specification.

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and accessibility
coefficients for three models using different specifications of the utility
equation, as well as the ten-minute walk buffer calculated by ParkScore. 
Note that the model called "Size and Distance" is precisely the model
labeled "SDEM" in Table \ref{tab:slx-sdem}. The controlling covariates are
suppressed for clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in the Appendix. Each of the three
utility-based specifications is significant at the $\alpha = 0.05$ level or lower,
and is of the hypothesized directionality. That is, residents living in tracts 
with increased utility-based accessiblity to green spaces show a significantly
higher physical activity participation rate. Taking the rough mean of the 2
estimated utility-based coefficients, moving from the least-accessible tract
($A_i = -3$) to the
most-accessible tract ($A_i = 3$) is expected to raise the physical activity rate
by `r 6 * .15` percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor park access
is anywhere between `r 6 * .01` and `r 6 * .3` percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the *obesity* rate in a tract, and physical activity becomes an independent covariate
alongside the controlling variables and the accessibility metrics.
Table \ref{tab:obesity-models} presents estimates of the effect of all three
utility-based accessibility measures as well as the 10-minute walk buffer. 
As before, the estimates of the other variables are available in the Appendix.
The correlation between physical activity rates and obesity is clear; for every
percentage point increase in physical activity rate, the expected obesity rate
drops by $0.43$ percentage points in all four accessiblity specifications.
The indirect effect of physical activity rates on obesity in neighboring zones
is inconsequential, supporting our methodological assertion that public health
variables are not spatially dependent, though the factors leading to them may be.

@@

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
accessibility and these metrics, previous literature has produced mixed
findings, perhaps owing to the range of variables modeled and the coarse spatial
scale of the analyses. Using New York City as a case study, we presented a
holistic and flexible measurement for park accessibility that compares the
continuous distance to all parks in the region, weighted against the size of the
park and its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive relationship
where the least park-accessible tracts have an expected physical activity
participation rate between `r 6 * .01` and `r 6 * .3` percentage points lower than the most
accessible tracts. We also found a suggestive though not sigificant relationship
between park accessiblity and expected obesity rates *in addition to* the effect
of physical activity participation.

>>>>

---
title: >
  Modeling the impacts of park access on health outcomes: a utility-based 
  accessibility approach
runninghead: Macfarlane \emph{et al}.
author:
  - name: Gregory S. Macfarlane
    num: 1
  - name: Nico Boyd
    num: 2
  - name: John E. Taylor
    num: 3
  - name: Kari Watkins
    num: 3
address:
  - code: 1
    address: Civil and Environmental Engineering Department, Brigham Young University 
  - code: 2
    address: Fehr and Peers, Los Angeles, California
  - code: 3
    address: School of Civil and Environmental Engineering, Georgia Institute of Technology
corrauth: "Gregory Macfarlane, 430 Engineering Building, Provo UT 84663"
email: gregmacfarlane@gmail.com
abstract: >
  Recent research has underscored the potential for public green spaces to influence individual and societal health outcomes, but empirical measurements of such influences have yielded mixed results to date, with particular disagreement surrounding how access to parks ought to be defined while controlling for alternate explanations. In this paper we apply a comprehensive measure of park accessibility drawn from random utility choice theory, which avoids arbitrary assertions of proximity while incorporating potentially numerous amenities and attributes of both the parks and the population. We apply this utility-based accessibility measure to correlate Census tract-level obesity and physical activity rate estimates from the Centers for Disease Control and Prevention's 500 Cities project with tract-level American Community Survey socioeconomic data in New York City, paired with geographic open space data from New York City. Controlling for the socioeconomic variables and spatially correlated error terms, we show a positive and significant relationship between park access and physical activity rates. The data also suggest a negative relationship between park access and obesity rates, beyond what is expected through physical activity and socioeconomics. In doing so, this research contributes a more comprehensive modeling approach for measuring the impact of park access on health, and may improve our understanding of the role  parks and access to them can serve in furthering public health objectives.
keywords: Accessibility; Green Space
classoption:
  - shortAfour
output: 
  rticles::sage_article:
    keep_md: yes
    keep_tex: yes
    includes: 
      in_header: preamble.tex
      after_body: appendix.tex
bibliography: library.bib
bibliographystyle: sageh.bst
---

@@

Applications of this isochrone-based framework are easily derived and relatively
common. 
ParkScore [@parkscore2019] is a sophisticated application of this approach where $D$ is a
carefully calculated 10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. @devries2003natural
showed the amount of green space within a 3-kilometer radius was positively associated
with a suite of self-reported health outcomes. Similarly, @devries2013streetscape
showed that streetscape greenery within 500 meters of a residence was associated with
perceived general health. Conversely, @carlson2012interactions found no relationship
between the health of senior citizens and the number of parks within 500 meters.

@@

In a destination choice framework, the tracts represent the "origins" and the
"destinations" are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and green spaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these distinct sub-facilities, we created 
variables for each containing park indicating the presence of sports courts,
playgrounds, and trails.

@@

Note that in the lag-$X$ specifications (SLX and SDEM) we exclude the accessibility 
component $\boldsymbol{A}$, that is we do *not* consider that the accessibility
in a neighboring zone will have an effect on physical activity or obesity rates.
In a practical sense, this is because the accessibility term $\boldsymbol{A}$ is itself spatially
determined.

@@

To determine which spatial effects specification was appropriate, we estimated
SLX and SDEM models of physical activity participation rates as a function of
tract covariates and  accessibility. These two models are presented in Table 
\ref{tab:slx-sdem}. With only a few exceptions, the SLX and SDEM coefficient
estimates for corresponding covariates lie within the 95\% confidence intervals
of the other model. A Hausman-style test [@Pace2008] on the coefficients 
assumes a null hypothesis that the coefficients are equivalent; this test 
produces $p$-value of only $`r slx_sdem_htest[2]`$. Though this statistic is
well beneath a common hypothesis test threshold of $\alpha = 0.05$, we
elect retain the null hypothesis that the coefficients are equivalent, as 
the differences are at no point practically different from each other. The consequences
of rejecting the null hypothesis in this case would be to suggest a model with 
spatial dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we adopt the
SDEM specification moving forward as the spatial lag of the errors is 
significant and the overall log likelihood of the SDEM is higher than the SLX 
specification.

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and accessibility
coefficients for three models using different specifications of the utility
equation, as well as the ten-minute walk buffer calculated by ParkScore. 
Note that the model called "Size and Distance" is precisely the model
labeled "SDEM" in Table \ref{tab:slx-sdem}. The controlling covariates are
suppressed for clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in Table \ref{tab:pa-fullmodels} of the Supplementary Material. Each of the three
utility-based specifications is significant at the $\alpha = 0.05$ level or lower,
and is of the hypothesized directionality. That is, residents living in tracts 
with increased utility-based accessibility to green spaces show a significantly
higher physical activity participation rate. Taking the rough mean of the 2
estimated utility-based coefficients, moving from the least-accessible tract
($A_i = -3$) to the
most-accessible tract ($A_i = 3$) is expected to raise the physical activity rate
by `r 6 * .15` percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor park access
is anywhere between `r 6 * .01` and `r 6 * .3` percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the *obesity* rate in a tract, and physical activity becomes an independent covariate
alongside the controlling variables and the accessibility metrics.
Table \ref{tab:obesity-models} presents estimates of the effect of all three
utility-based accessibility measures as well as the 10-minute walk buffer. 
As before, the estimates of the other variables are available in Table 
\ref{tab:ob-fullmodels} of the Supplementary Material.
The correlation between physical activity rates and obesity is clear; for every
percentage point increase in physical activity rate, the expected obesity rate
drops by $0.43$ percentage points in all four accessibility specifications.
The indirect effect of physical activity rates on obesity in neighboring zones
is inconsequential, supporting our methodological assertion that public health
variables are not spatially dependent, though the factors leading to them may be.

@@

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
accessibility and these metrics, previous literature has produced mixed
findings, perhaps owing to the range of variables modeled and the coarse spatial
scale of the analyses. Using New York City as a case study, we presented a
holistic and flexible measurement for park accessibility that compares the
continuous distance to all parks in the region, weighted against the size of the
park and its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive relationship
where the least park-accessible tracts have an expected physical activity
participation rate between `r 6 * .01` and `r 6 * .3` percentage points lower than the most
accessible tracts. We also found a suggestive though not significant relationship
between park accessibility and expected obesity rates *in addition to* the effect
of physical activity participation.

### 2020-09-22 manuscript/manuscript_complete_source.tex
Submit first revisions

@@

\author{Gregory Macfarlane\affilnum{1}, Nico Boyd\affilnum{2}, John E. Taylor\affilnum{3}, Kari Watkins\affilnum{3}}

@@

Applications of this isochrone-based framework are easily derived and
relatively common. ParkScore \citep{parkscore2019} is a sophisticated
application of this approach where \(D\) is a carefully calculated
10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. \citet{devries2003natural}
showed the amount of greenspace within a 3-kilometer radius was
positively associated with a suite of self-reported health outcomes.
Similarly, \citet{devries2013streetscape} showed that streetscape
greenery within 500 meters of a residence was associated with perceived
general health. Conversely, \citet{carlson2012interactions} found no
relationship between the health of senior citizens and the number of
parks within 500 meters.

@@

In a destination choice framework, the tracts represent the ``origins''
and the ``destinations'' are parks and green spaces in New York City. We
retrieved a polygons layer of public parks and greenspaces within New
York City's municipal boundaries and checked it for accuracy and
relevance \citep{nycparks}. Upon inspection, we removed several
facilities that do not qualify as publicly accessible green space, such
as Yankee Stadium, Citi Field, and their surrounding parking lots. We
also removed parks of less than half an acre in size, as these appear to
be predominantly planted medians rather than legitimate public green
space. We consolidated individual geographic polygons comprising a
single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields
included within larger park facilities. Instead of these distinct
sub-facilities, we created variables for each containing park indicating
the presence of sports courts, playgrounds, and trails.

@@

Note that in the lag-\(X\) specifications (SLX and SDEM) we exclude the
accessiblity component \(\boldsymbol{A}\), that is we do \emph{not}
consider that the accessibility in a neighboring zone will have an
effect on physical activity or obesity rates. In a practical sense, this
is because the accessibility term \(\boldsymbol{A}\) is itself spatially
determined.

@@

To determine which spatial effects specification was appropriate, we
estimated SLX and SDEM models of physical activity participation rates
as a function of tract covariates and accessibility. These two models
are presented in Table \ref{tab:slx-sdem}. With only a few exceptions,
the SLX and SDEM coefficient estimates for corresponding covariates lie
within the 95\% confidence intervals of the other model. A Hausman-style
test \citep{Pace2008} on the coefficients assumes a null hypothesis that
the coefficients are equivalent; this test produces \(p\)-value of only
\(1.307e-23\). Though this statistic is well beneath a common hypothesis
test threshold of \(\alpha = 0.05\), we elect retain the null hypothesis
that the coefficients are equivalent, as the differences are at no point
practically different from each other. The consequences of rejecting the
null hypothesis in this case would be to suggest a model with spatial
dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we
adopt the SDEM specification moving forward as the spatial lag of the
errors is significant and the overal log likelihood of the SDEM is
higher than the SLX specification.

\begin{table*}
\caption{Comparison of SLX and SDEM Coefficients}
\label{tab:slx-sdem}
\begin{tabular}{l d{2.4} d{2.4} d{2.4} c d{2.4} d{2.4} d{2.4} }
\toprule
 & \multicolumn{3}{c}{SLX} & & \multicolumn{3}{c}{SDEM} \\
\cmidrule{2-4}\cmidrule{6-8}
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.} & 
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.}\\
\midrule
% latex table generated in R 4.0.0 by xtable 1.8-4 package
% Thu Jun  4 16:31:52 2020
 (Intercept) & -15.8424 & -26.8735 & -4.8112 &   & -24.9740 & -41.8930 & -8.0549 \\ 
  log(Density) & 0.0484 & -0.1314 & 0.2282 &   & 0.1775 & 0.0384 & 0.3166 \\ 
  log(Income) & 5.8210 & 5.2977 & 6.3443 &   & 5.9985 & 5.5746 & 6.4224 \\ 
  Fulltime & 0.1150 & 0.0936 & 0.1365 &   & 0.1275 & 0.1090 & 0.1460 \\ 
  College-educated & 0.0012 & -0.0272 & 0.0297 &   & 0.0076 & -0.0158 & 0.0309 \\ 
  Single Adults & -0.0433 & -0.0623 & -0.0244 &   & -0.0362 & -0.0526 & -0.0198 \\ 
  Youth (0-17) & -0.1511 & -0.1830 & -0.1192 &   & -0.1310 & -0.1576 & -0.1045 \\ 
  Young adults (18-34) & 0.0257 & 0.0008 & 0.0505 &   & 0.0295 & 0.0094 & 0.0496 \\ 
  Seniors (65+) & 0.0137 & -0.0185 & 0.0459 &   & 0.0403 & 0.0130 & 0.0677 \\ 
  Black population share & -0.0465 & -0.0584 & -0.0346 &   & -0.0499 & -0.0586 & -0.0411 \\ 
  Asian population share & -0.0686 & -0.0832 & -0.0541 &   & -0.0767 & -0.0873 & -0.0662 \\ 
  Hispanic population share & -0.1017 & -0.1140 & -0.0894 &   & -0.1025 & -0.1116 & -0.0934 \\ 
  Other Minorities & 0.0368 & -0.0643 & 0.1378 &   & 0.0037 & -0.0849 & 0.0923 \\ 
  $\gamma$: log(Density) & 1.2594 & 0.9413 & 1.5774 &   & 1.0985 & 0.7155 & 1.4815 \\ 
  $\gamma$: log(Income) & 2.3930 & 1.5476 & 3.2383 &   & 1.9742 & 0.9465 & 3.0020 \\ 
  $\gamma$: Fulltime & -0.1398 & -0.1804 & -0.0991 &   & -0.0397 & -0.0869 & 0.0076 \\ 
  $\gamma$: College-educated & -0.1312 & -0.1806 & -0.0818 &   & -0.0710 & -0.1305 & -0.0115 \\ 
  $\gamma$: Single Adults & -0.0480 & -0.0821 & -0.0140 &   & -0.0082 & -0.0504 & 0.0339 \\ 
  $\gamma$: Youth (0-17) & -0.1031 & -0.1618 & -0.0445 &   & -0.0108 & -0.0798 & 0.0583 \\ 
  $\gamma$: Young adults (18-34) & 0.1191 & 0.0762 & 0.1619 &   & 0.0842 & 0.0335 & 0.1349 \\ 
  $\gamma$: Seniors (65+) & 0.0108 & -0.0479 & 0.0695 &   & 0.0590 & -0.0102 & 0.1282 \\ 
  $\gamma$: Black population share & 0.0096 & -0.0054 & 0.0247 &   & 0.0010 & -0.0151 & 0.0172 \\ 
  $\gamma$: Asian population share & -0.0661 & -0.0840 & -0.0483 &   & -0.0397 & -0.0583 & -0.0210 \\ 
  $\gamma$: Hispanic population share & -0.0060 & -0.0216 & 0.0097 &   & -0.0010 & -0.0182 & 0.0162 \\ 
  $\gamma$: Other Minorities & 0.1868 & -0.0508 & 0.4244 &   & 0.0726 & -0.1845 & 0.3297 \\ 
  Accessibility & 0.5430 & 0.4195 & 0.6666 &   & 0.1663 & 0.0277 & 0.3048 \\ 
  $\lambda$: spatial correlation &  &  &  &   & 0.6905 & 0.6513 & 0.7296 \\

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and
accessibility coefficients for three models using different
specifications of the utility equation, as well as the ten-minute walk
buffer calculated by ParkScore. Note that the model called ``Size and
Distance'' is precisely the model labeled ``SDEM'' in Table
\ref{tab:slx-sdem}. The controlling covariates are suppressed for
clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in the Appendix. Each of the
three utility-based specifications is significant at the
\(\alpha = 0.05\) level or lower, and is of the hypothesized
directionality. That is, residents living in tracts with increased
utility-based accessiblity to green spaces show a significantly higher
physical activity participation rate. Taking the rough mean of the 2
estimated utility-based coefficients, moving from the least-accessible
tract (\(A_i = -3\)) to the most-accessible tract (\(A_i = 3\)) is
expected to raise the physical activity rate by 0.9 percentage points.
With the understanding of the full 95\% confidence interval, we could
suggest that the true effect of excellent versus poor park access is
anywhere between 0.06 and 1.8 percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the \emph{obesity} rate in a tract, and physical activity becomes an
independent covariate alongside the controlling variables and the
accessibility metrics. Table \ref{tab:obesity-models} presents estimates
of the effect of all three utility-based accessibility measures as well
as the 10-minute walk buffer. As before, the estimates of the other
variables are available in the Appendix. The correlation between
physical activity rates and obesity is clear; for every percentage point
increase in physical activity rate, the expected obesity rate drops by
\(0.43\) percentage points in all four accessiblity specifications. The
indirect effect of physical activity rates on obesity in neighboring
zones is inconsequential, supporting our methodological assertion that
public health variables are not spatially dependent, though the factors
leading to them may be.

@@

Increased physical activity and decreased obesity rates are critical
measures of improvement in public health. Although many have theorized
the link between park accessibility and these metrics, previous
literature has produced mixed findings, perhaps owing to the range of
variables modeled and the coarse spatial scale of the analyses. Using
New York City as a case study, we presented a holistic and flexible
measurement for park accessibility that compares the continuous distance
to all parks in the region, weighted against the size of the park and
its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive
relationship where the least park-accessible tracts have an expected
physical activity participation rate between 0.06 and 1.8 percentage
points lower than the most accessible tracts. We also found a suggestive
though not sigificant relationship between park accessiblity and
expected obesity rates \emph{in addition to} the effect of physical
activity participation.

@@

\hypertarget{appendix}{%
\subsection{Appendix}\label{appendix}}

In this appendix we present the complete estimation results for the
models relating different definitions of access to physical activity (in
Table \ref{tab:pa-fullmodels}) and to obesity (in Table
\ref{tab:ob-fullmodels}). In each case we also present a base model with
no accessibility statistics for comparison.

>>>>

\renewcommand{\appendixname}{Supplementary Material}

@@

\author{Gregory S. Macfarlane\affilnum{1}, Nico Boyd\affilnum{2}, John E. Taylor\affilnum{3}, Kari Watkins\affilnum{3}}

@@

Applications of this isochrone-based framework are easily derived and
relatively common. ParkScore \citep{parkscore2019} is a sophisticated
application of this approach where \(D\) is a carefully calculated
10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. \citet{devries2003natural}
showed the amount of green space within a 3-kilometer radius was
positively associated with a suite of self-reported health outcomes.
Similarly, \citet{devries2013streetscape} showed that streetscape
greenery within 500 meters of a residence was associated with perceived
general health. Conversely, \citet{carlson2012interactions} found no
relationship between the health of senior citizens and the number of
parks within 500 meters.

@@

In a destination choice framework, the tracts represent the ``origins''
and the ``destinations'' are parks and green spaces in New York City. We
retrieved a polygons layer of public parks and green spaces within New
York City's municipal boundaries and checked it for accuracy and
relevance \citep{nycparks}. Upon inspection, we removed several
facilities that do not qualify as publicly accessible green space, such
as Yankee Stadium, Citi Field, and their surrounding parking lots. We
also removed parks of less than half an acre in size, as these appear to
be predominantly planted medians rather than legitimate public green
space. We consolidated individual geographic polygons comprising a
single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields
included within larger park facilities. Instead of these distinct
sub-facilities, we created variables for each containing park indicating
the presence of sports courts, playgrounds, and trails.

@@

Note that in the lag-\(X\) specifications (SLX and SDEM) we exclude the
accessibility component \(\boldsymbol{A}\), that is we do \emph{not}
consider that the accessibility in a neighboring zone will have an
effect on physical activity or obesity rates. In a practical sense, this
is because the accessibility term \(\boldsymbol{A}\) is itself spatially
determined.

@@

To determine which spatial effects specification was appropriate, we
estimated SLX and SDEM models of physical activity participation rates
as a function of tract covariates and accessibility. These two models
are presented in Table \ref{tab:slx-sdem}. With only a few exceptions,
the SLX and SDEM coefficient estimates for corresponding covariates lie
within the 95\% confidence intervals of the other model. A Hausman-style
test \citep{Pace2008} on the coefficients assumes a null hypothesis that
the coefficients are equivalent; this test produces \(p\)-value of only
\(1.307e-23\). Though this statistic is well beneath a common hypothesis
test threshold of \(\alpha = 0.05\), we elect retain the null hypothesis
that the coefficients are equivalent, as the differences are at no point
practically different from each other. The consequences of rejecting the
null hypothesis in this case would be to suggest a model with spatial
dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we
adopt the SDEM specification moving forward as the spatial lag of the
errors is significant and the overall log likelihood of the SDEM is
higher than the SLX specification.

\begin{table*}
\caption{Comparison of SLX and SDEM Coefficients}
\label{tab:slx-sdem}
\begin{tabular}{l d{2.4} d{2.4} d{2.4} c d{2.4} d{2.4} d{2.4} }
\toprule
 & \multicolumn{3}{c}{SLX} & & \multicolumn{3}{c}{SDEM} \\
\cmidrule{2-4}\cmidrule{6-8}
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.} & 
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.}\\
\midrule
% latex table generated in R 4.0.0 by xtable 1.8-4 package
% Tue Jul  7 09:56:24 2020
 (Intercept) & -15.8424 & -26.8735 & -4.8112 &   & -24.9740 & -41.8930 & -8.0549 \\ 
  log(Density) & 0.0484 & -0.1314 & 0.2282 &   & 0.1775 & 0.0384 & 0.3166 \\ 
  log(Income) & 5.8210 & 5.2977 & 6.3443 &   & 5.9985 & 5.5746 & 6.4224 \\ 
  Fulltime & 0.1150 & 0.0936 & 0.1365 &   & 0.1275 & 0.1090 & 0.1460 \\ 
  College-educated & 0.0012 & -0.0272 & 0.0297 &   & 0.0076 & -0.0158 & 0.0309 \\ 
  Single Adults & -0.0433 & -0.0623 & -0.0244 &   & -0.0362 & -0.0526 & -0.0198 \\ 
  Youth (0-17) & -0.1511 & -0.1830 & -0.1192 &   & -0.1310 & -0.1576 & -0.1045 \\ 
  Young adults (18-34) & 0.0257 & 0.0008 & 0.0505 &   & 0.0295 & 0.0094 & 0.0496 \\ 
  Seniors (65+) & 0.0137 & -0.0185 & 0.0459 &   & 0.0403 & 0.0130 & 0.0677 \\ 
  Black population share & -0.0465 & -0.0584 & -0.0346 &   & -0.0499 & -0.0586 & -0.0411 \\ 
  Asian population share & -0.0686 & -0.0832 & -0.0541 &   & -0.0767 & -0.0873 & -0.0662 \\ 
  Hispanic population share & -0.1017 & -0.1140 & -0.0894 &   & -0.1025 & -0.1116 & -0.0934 \\ 
  Other Minorities & 0.0368 & -0.0643 & 0.1378 &   & 0.0037 & -0.0849 & 0.0923 \\ 
  $\gamma$: log(Density) & 1.2594 & 0.9413 & 1.5774 &   & 1.0985 & 0.7155 & 1.4815 \\ 
  $\gamma$: log(Income) & 2.3930 & 1.5476 & 3.2383 &   & 1.9742 & 0.9465 & 3.0020 \\ 
  $\gamma$: Fulltime & -0.1398 & -0.1804 & -0.0991 &   & -0.0397 & -0.0869 & 0.0076 \\ 
  $\gamma$: College-educated & -0.1312 & -0.1806 & -0.0818 &   & -0.0710 & -0.1305 & -0.0115 \\ 
  $\gamma$: Single Adults & -0.0480 & -0.0821 & -0.0140 &   & -0.0082 & -0.0504 & 0.0339 \\ 
  $\gamma$: Youth (0-17) & -0.1031 & -0.1618 & -0.0445 &   & -0.0108 & -0.0798 & 0.0583 \\ 
  $\gamma$: Young adults (18-34) & 0.1191 & 0.0762 & 0.1619 &   & 0.0842 & 0.0335 & 0.1349 \\ 
  $\gamma$: Seniors (65+) & 0.0108 & -0.0479 & 0.0695 &   & 0.0590 & -0.0102 & 0.1282 \\ 
  $\gamma$: Black population share & 0.0096 & -0.0054 & 0.0247 &   & 0.0010 & -0.0151 & 0.0172 \\ 
  $\gamma$: Asian population share & -0.0661 & -0.0840 & -0.0483 &   & -0.0397 & -0.0583 & -0.0210 \\ 
  $\gamma$: Hispanic population share & -0.0060 & -0.0216 & 0.0097 &   & -0.0010 & -0.0182 & 0.0162 \\ 
  $\gamma$: Other Minorities & 0.1868 & -0.0508 & 0.4244 &   & 0.0726 & -0.1845 & 0.3297 \\ 
  Accessibility & 0.5430 & 0.4195 & 0.6666 &   & 0.1663 & 0.0277 & 0.3048 \\ 
  $\lambda$: spatial correlation &  &  &  &   & 0.6905 & 0.6513 & 0.7296 \\

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and
accessibility coefficients for three models using different
specifications of the utility equation, as well as the ten-minute walk
buffer calculated by ParkScore. Note that the model called ``Size and
Distance'' is precisely the model labeled ``SDEM'' in Table
\ref{tab:slx-sdem}. The controlling covariates are suppressed for
clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in Table \ref{tab:pa-fullmodels}
of the Supplementary Material. Each of the three utility-based
specifications is significant at the \(\alpha = 0.05\) level or lower,
and is of the hypothesized directionality. That is, residents living in
tracts with increased utility-based accessibility to green spaces show a
significantly higher physical activity participation rate. Taking the
rough mean of the 2 estimated utility-based coefficients, moving from
the least-accessible tract (\(A_i = -3\)) to the most-accessible tract
(\(A_i = 3\)) is expected to raise the physical activity rate by 0.9
percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor
park access is anywhere between 0.06 and 1.8 percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the \emph{obesity} rate in a tract, and physical activity becomes an
independent covariate alongside the controlling variables and the
accessibility metrics. Table \ref{tab:obesity-models} presents estimates
of the effect of all three utility-based accessibility measures as well
as the 10-minute walk buffer. As before, the estimates of the other
variables are available in Table \ref{tab:ob-fullmodels} of the
Supplementary Material. The correlation between physical activity rates
and obesity is clear; for every percentage point increase in physical
activity rate, the expected obesity rate drops by \(0.43\) percentage
points in all four accessibility specifications. The indirect effect of
physical activity rates on obesity in neighboring zones is
inconsequential, supporting our methodological assertion that public
health variables are not spatially dependent, though the factors leading
to them may be.

@@

Increased physical activity and decreased obesity rates are critical
measures of improvement in public health. Although many have theorized
the link between park accessibility and these metrics, previous
literature has produced mixed findings, perhaps owing to the range of
variables modeled and the coarse spatial scale of the analyses. Using
New York City as a case study, we presented a holistic and flexible
measurement for park accessibility that compares the continuous distance
to all parks in the region, weighted against the size of the park and
its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive
relationship where the least park-accessible tracts have an expected
physical activity participation rate between 0.06 and 1.8 percentage
points lower than the most accessible tracts. We also found a suggestive
though not significant relationship between park accessibility and
expected obesity rates \emph{in addition to} the effect of physical
activity participation.

@@

\hypertarget{appendix}{%
\subsection{Supplementary Material}\label{appendix}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}

In this supplementary material we present the complete estimation results for the
models relating different definitions of access to physical activity (in
Table \ref{tab:pa-fullmodels}) and to obesity (in Table
\ref{tab:ob-fullmodels}). In each case we also present a base model with
no accessibility statistics for comparison.
